Tracking p-adic precision: examples

### 2014-08-21 examples.tex
Typos
Correction of some typos, plus some small details added.

Let $U$, $V$ and $W$ be three open subsets is some $K$-Banach spaces
$E$, $F$ and $G$ respectively. We assume that $0 \in U$, $0 \in V$ and
consider two locally analytic functions $f : U \to V$ and $g : V \to W$ 
two locally analytic functions around $0$. The composition $h = g \circ
f$ is then locally analytic around $0$ as well. Let us write the
analytic expansion of $f$, $g$ and $h$ as follows:
$$f = \sum_{n \geq 0} f_n, \quad 
g = \sum_{n \geq 0} g_n, \quad
h = \sum_{n \geq 0} h_n, \quad$$
where $f_n$, $g_n$ and $h_n$ are the restrictions to the diagonal of 
some $n$-linear forms $F_n$, $G_n$ and $H_n$ respectively. The aim of
this subsection is to prove the following proposition.

@@

\begin{proof}
It is enough to prove that:
\begin{equation}
\label{eq:normpolar}
\Big\Vert \binom m {k_1 \,\, k_2 \,\, \cdots \,\, k_\ell} \cdot
\varphi(x_1, \ldots, x_1, \ldots, x_\ell, \ldots,
x_\ell) \Big\Vert \leq \Vert \psi \Vert
\end{equation}
provided that all the $x_i$'s have norm at most $1$. We proceed by 
induction on 
$\ell$. If $\ell = 1$, Eq.~\eqref{eq:normpolar} follows directly from
the definition of $\Vert \psi \Vert$. 
We now pick $(\ell+1)$ integers $k_1, \ldots, k_{\ell+1}$ whose sum 
equals $m$ together with $(\ell+1)$ elements $x_1, \ldots, x_{\ell+1}$
lying in the unit ball of $\mathcal E$.
We also consider a new variable $\lambda$ varying in
$\O_K$. We set $x'_i = x_i$, $k'_i = k_i$ when $i < \ell$ and $x'_\ell 
= x_\ell + \lambda x_{\ell+1}$, $k_\ell = k_\ell + k_{\ell+1}$. By the 
induction hypothesis, we know that the inequality:
$$\Big\Vert \binom s {k'_1 \,\, \cdots \,\, k'_\ell} \cdot
\varphi(x'_1, \ldots, x'_1, \ldots, x'_\ell, \ldots, x'_\ell) \Big\Vert
\leq \Vert \psi \Vert$$
holds for all $\lambda \in K$. Furthermore, LHS of the above inequality
is a polynomial $P(\lambda)$ of degree $k'_\ell$ whose coefficient in 
$\lambda^j$ is:
$$\binom s {k'_1 \,\, \cdots \,\, k'_\ell} \cdot
\binom {k'_\ell} {j} \cdot
\varphi(\underline x_j) = 
\binom s {k_1 \,\, \cdots \,\, k_{\ell-1} \,\, j} \cdot
\varphi(\underline x_j)$$
with
$$\underline x_j = (x_1, \ldots, x_1, \ldots, x_{\ell+1}, \ldots, 
x_{\ell+1})$$
where $x_i$ is repeated $k_i$ times if $i < \ell$ and $x_\ell$ 
(resp. $x_{\ell+1}$) is repeated $j$ times (resp. $k'_\ell - j$ times).
Since $\Vert P(\lambda) \Vert \leq \Vert \psi \Vert$ for all $\lambda$ in 
the unit ball, the norm of all its coefficients must be at most $\Vert \psi
\Vert$ as well. Looking at the coefficient in $\lambda^{k_i}$, we find
Eq.~\eqref{eq:normpolar} for the families $(k_1, \ldots, k_{\ell+1})$ 
and $(x_1, \ldots, x_{\ell+1})$. The induction follows.
\end{proof}

@@

We now study a slight generalization of the differential equation 
\eqref{eq:diffequa}, which is:
\begin{equation}
\label{eq:diffequah}
f' = g \circ (f, h).
\end{equation}
Here $g$ and $h$ are known locally analytic functions and $f$ is the 
unknown, which is assumed to be locally analytic as well. The function
$f$ goes from $U$ to $V$, two open subsets in $K$-Banach spaces $E$ 
and $F$ respectively. The function $h$ goes from $U$ to $W$ where $W$
is again an open subset in a $K$-Banach space $G$. Consequently, $g$
goes from $V \times W$ to the vector space $\mathcal L(E,F)$ gathering
all continuous linear applications $E \to F$.
In what follows, we always assume that $V$ and $W$ contains the origin, 
$f(0) = 0$, $h(0) = 0$ and $g(0) \neq 0$. There assumptions are armless 
because (1)~we can always shift $f$ and $h$ (and $g$ accordingly) so 
that they both vanish at $0$ and (2)~in order to apply the theory of 
\cite{padicprec}, the derivative $f'(0)$ needs to be surjective and 
therefore \emph{a fortiori} nonzero.

@@

We now pick two real numbers $A$ and $B$ satisfying \eqref{eq:condAB}.
Set $d = \Lambda(h)(a)$. Going back to the definitions, we remark that
the above equality implies:
\begin{equation}
\label{eq:boundhn}
\Vert h_n \Vert \leq e^{- a n + d},
  \quad \forall n \in \N.
\end{equation}
In the same way, from Eq.~\eqref{eq:condAB}, we get:
\begin{equation}
\label{eq:boundgm}
\Vert g_m \Vert \leq e^{- \max(b,d)\cdot m + b - a},
  \quad \forall m \in \N.
\end{equation}
We are now ready to prove $u_r \leq e^{-ar + b}$ by induction on $r$.
When $r = 0$, it is obvious because $u_0$ vanishes. Otherwise, it
follows from \eqref{eq:boundurrec}, \eqref{eq:boundhn}, \eqref{eq:boundgm}, 
and the induction hypothesis that:
\begin{align*}
u_{r+1} 
& \leq \sup_{m, (n_i)}
    e^{ -\max(b,d)\cdot m + b - a + \sum_{i=1}^m (-a n_i + \max(b,d))} \\
& = e^{ b - a - a r } = e^{ -a (r+1) + b}
\end{align*}
and the induction goes.
Finally, keeping in mind the definition of $u_r$, from $u_r \leq 
e^{-a r + b}$, we obtain:
$$\Vert f_r \Vert \leq u_r \cdot \Vert r! \Vert^{-1} \leq
e^{-(a - \alpha) r + b}$$
which means that $b \geq \Lambda(f)(a - \alpha)$.
\end{proof}

@@

Figure \ref{fig:area} illustrates Corollary \ref{cor:boundLambdaf}. The 
blue plain line represents the graph of the function $\Lambda_f$. A 
quick computation shows that, on a neighborhood of ${-}\infty$, this 
function is given by:
$$\Lambda_f(x) = x + \alpha + \mu$$
where $\mu$ is the value that $\Lambda_g$ takes on the interval 
$]{-}\infty, \nu]$. Corollary \ref{cor:boundLambdaf} says that the
graph of $\Lambda(f)$ lies below the plain blue line. We remark
moreover that the Taylor expansion of $f(x)$ starts with the term
$g(0)(x)$. Hence, on a neighborhood on ${-}\infty$, we have 
$\Lambda(f)(x) = x + \log \Vert g(0) \Vert$. Using convexity, we 
get:
$$\Lambda(f)(x) \geq x + \log \Vert g(0) \Vert, 
  \quad \forall x \in \R.$$
In other words, the graph of $\Lambda(f)$ lies above the brown line.
Furthermore, we know that the slopes of $\Lambda(f)$ are all integral
because $f$ is locally analytic. Hence, $\Lambda(f)$ cannot lie above
the dashed blue line defined as the line of slope $2$ passing through
the first break point of the blue plain line --- which has coordinate 
$(y_0 - \alpha - \mu, y_0)$ with $y_0 = \min(\Lambda_h^{-1}(\nu) + \mu, 
\nu)$. As a conclusion, we have proved that the graph of $\Lambda(f)$ 
must coincide with the brown line until it meets the dashed blue line 
and then has to stay in the green area.

@@

We believe that a good way to opmitize $\lambda$ is to choose it so that 
the two quantities in the minima in Proposition \ref{prop:boundLambdaf2} 
agree. With the expression we got just above for $\Lambda_h$, our 
heuristic yields $\log \lambda = \mu$, \emph{i.e.} $\lambda = e^\mu$.
Making this choice, the norms of $f(x)$ and $h(x)$ are comparable --- 
at least, on the domain we are interested in --- and we do not loose so 
much when we write that the norm of the couple $(f(x), h(x))$ is the 
maximum of those of $f(x)$ and $h(x)$.

>>>>

Let $U$, $V$ and $W$ be three open subsets is some $K$-Banach spaces
$E$, $F$ and $G$ respectively. We assume that $0 \in U$, $0 \in V$ and
consider two locally analytic functions $f : U \to V$ and $g : V \to W$ 
two locally analytic functions around $0$, with $f(0)=0.$ The composition $h = g \circ
f$ is then locally analytic around $0$ as well. Let us write the
analytic expansion of $f$, $g$ and $h$ as follows:
$$f = \sum_{n \geq 0} f_n, \quad 
g = \sum_{n \geq 0} g_n, \quad
h = \sum_{n \geq 0} h_n, \quad$$
where $f_n$, $g_n$ and $h_n$ are the restrictions to the diagonal of 
some $n$-linear forms $F_n$, $G_n$ and $H_n$ respectively. The aim of
this subsection is to prove the following proposition.

@@

\begin{proof}
It is enough to prove that:
\begin{equation}
\label{eq:normpolar}
\Big\Vert \binom m {k_1 \,\, k_2 \,\, \cdots \,\, k_\ell} \cdot
\varphi(x_1, \ldots, x_1, \ldots, x_\ell, \ldots,
x_\ell) \Big\Vert \leq \Vert \psi \Vert
\end{equation}
provided that all the $x_i$'s have norm at most $1$. We proceed by 
induction on 
$\ell$. If $\ell = 1$, Eq.~\eqref{eq:normpolar} follows directly from
the definition of $\Vert \psi \Vert$. 
We now pick $(\ell+1)$ integers $k_1, \ldots, k_{\ell+1}$ whose sum 
equals $m$ together with $(\ell+1)$ elements $x_1, \ldots, x_{\ell+1}$
lying in the unit ball of $\mathcal E$.
We also consider a new variable $\lambda$ varying in
$\O_K$. We set $x'_i = x_i$, $k'_i = k_i$ when $i < \ell$ and $x'_\ell 
= x_\ell + \lambda x_{\ell+1}$, $k'_\ell = k_\ell + k_{\ell+1}$. By the 
induction hypothesis, we know that the inequality:
$$\Big\Vert \binom s {k'_1 \,\, \cdots \,\, k'_\ell} \cdot
\varphi(x'_1, \ldots, x'_1, \ldots, x'_\ell, \ldots, x'_\ell) \Big\Vert
\leq \Vert \psi \Vert$$
holds for all $\lambda \in K$. Furthermore, LHS of the above inequality
is a polynomial $P(\lambda)$ of degree $k'_\ell$ whose coefficient in 
$\lambda^j$ is:
$$\binom s {k'_1 \,\, \cdots \,\, k'_\ell} \cdot
\binom {k'_\ell} {j} \cdot
\varphi(\underline x_j) = 
\binom s {k_1 \,\, \cdots \,\, k_{\ell-1} \,\, j} \cdot
\varphi(\underline x_j)$$
with
$$\underline x_j = (x_1, \ldots, x_1, \ldots, x_{\ell+1}, \ldots, 
x_{\ell+1})$$
where $x_i$ is repeated $k_i$ times if $i < \ell$ and $x_\ell$ 
(resp. $x_{\ell+1}$) is repeated $j$ times (resp. $k'_\ell - j$ times).
Since $\Vert P(\lambda) \Vert \leq \Vert \psi \Vert$ for all $\lambda$ in 
the unit ball, the norm of all its coefficients must be at most $\Vert \psi
\Vert$ as well. Looking at the coefficient in $\lambda^{k_{l+1}}$, we find
Eq.~\eqref{eq:normpolar} for the families $(k_1, \ldots, k_{\ell+1})$ 
and $(x_1, \ldots, x_{\ell+1})$. The induction follows.
\end{proof}

@@

We now study a slight generalization of the differential equation 
\eqref{eq:diffequa}, which is:
\begin{equation}
\label{eq:diffequah}
f' = g \circ (f, h).
\end{equation}
Here $g$ and $h$ are known locally analytic functions and $f$ is the 
unknown, which is assumed to be locally analytic as well. The function
$f$ goes from $U$ to $V$, two open subsets in $K$-Banach spaces $E$ 
and $F$ respectively. The function $h$ goes from $U$ to $W$ where $W$
is again an open subset in a $K$-Banach space $G$. Consequently, $g$
goes from $V \times W$ to the vector space $\mathcal L(E,F)$ gathering
all continuous linear applications $E \to F$.
In what follows, we always assume that $V$ and $W$ contains the origin, 
$f(0) = 0$, $h(0) = 0$ and $g(0) \neq 0$. There assumptions are harmless 
because (1)~we can always shift $f$ and $h$ (and $g$ accordingly) so 
that they both vanish at $0$ and (2)~in order to apply the theory of 
\cite{padicprec}, the derivative $f'(0)$ needs to be surjective and 
therefore \emph{a fortiori} nonzero.

@@

We now pick two real numbers $a$ and $b$ satisfying \eqref{eq:condAB}.
Set $d = \Lambda(h)(a)$. Going back to the definitions of $\Lambda (h)$ and Legendre transform, we remark that
the above equality implies:
\begin{equation}
\label{eq:boundhn}
\Vert h_n \Vert \leq e^{- a n + d},
  \quad \forall n \in \N.
\end{equation}
In the same way, from Eq.~\eqref{eq:condAB}, we get:
\begin{equation}
\label{eq:boundgm}
\Vert g_m \Vert \leq e^{- \max(b,d)\cdot m + b - a},
  \quad \forall m \in \N.
\end{equation}
We are now ready to prove $u_r \leq e^{-ar + b}$ by induction on $r$.
When $r = 0$, it is obvious because $u_0$ vanishes. Otherwise, it
follows from \eqref{eq:boundurrec}, \eqref{eq:boundhn}, \eqref{eq:boundgm}, 
and the induction hypothesis that:
\begin{align*}
u_{r+1} 
& \leq \sup_{m, (n_i)}
    e^{ -\max(b,d)\cdot m + b - a + \sum_{i=1}^m (-a n_i + \max(b,d))} \\
& = e^{ b - a - a r } = e^{ -a (r+1) + b}
\end{align*}
and the induction goes.
Finally, keeping in mind the definition of $u_r$, from $u_r \leq 
e^{-a r + b}$, we obtain:
$$\Vert f_r \Vert \leq u_r \cdot \Vert r! \Vert^{-1} \leq
e^{-(a - \alpha) r + b}$$
which means that $b \geq \Lambda(f)(a - \alpha)$.
\end{proof}

@@

Figure \ref{fig:area} illustrates Corollary \ref{cor:boundLambdaf}. The 
blue plain line represents the graph of the function $\Lambda_f$. A 
quick computation shows that, on a neighborhood of ${-}\infty$, this 
function is given by:
$$\Lambda_f(x) = x + \alpha + \mu$$
where $\mu$ is the value that $\Lambda_g$ takes on the interval 
$]{-}\infty, \nu]$. Corollary \ref{cor:boundLambdaf} says that the
graph of $\Lambda(f)$ lies below the plain blue line. We remark
moreover that the Taylor expansion of $f(x)$ starts with the term
$g(0) x$. Hence, on a neighborhood on ${-}\infty$, we have 
$\Lambda(f)(x) = x + \log \Vert g(0) \Vert$. Using convexity, we 
get:
$$\Lambda(f)(x) \geq x + \log \Vert g(0) \Vert, 
  \quad \forall x \in \R.$$
In other words, the graph of $\Lambda(f)$ lies above the brown line.
Furthermore, we know that the slopes of $\Lambda(f)$ are all integral
because $f$ is locally analytic. Hence, $\Lambda(f)$ cannot lie above
the dashed blue line defined as the line of slope $2$ passing through
the first break point of the blue plain line --- which has coordinate 
$(y_0 - \alpha - \mu, y_0)$ with $y_0 = \min(\Lambda_h^{-1}(\nu) + \mu, 
\nu)$. As a conclusion, we have proved that the graph of $\Lambda(f)$ 
must coincide with the brown line until it meets the dashed blue line 
and then has to stay in the green area.

@@

We believe that a good way to optimize $\lambda$ is to choose it so that 
the two quantities in the minima in Proposition \ref{prop:boundLambdaf2} 
agree. With the expression we got just above for $\Lambda_h$, our 
heuristic yields $\log \lambda = \mu$, \emph{i.e.} $\lambda = e^\mu$.
Making this choice, the norms of $f(x)$ and $h(x)$ are comparable --- 
at least, on the domain we are interested in --- and we do not loose so 
much when we write that the norm of the couple $(f(x), h(x))$ is the 
maximum of those of $f(x)$ and $h(x)$.
